%==============================================================================
% Notes for CSCI 2951-F (Fall 2015)
%==============================================================================
\newcommand{\thishw}{\bf Complexity Reading Group 11/19/2015}
\newcommand{\myname}{David Abel}

%==============================================================================
% Mini Preamble.
%==============================================================================

\documentclass[11pt]{article} % 10pt article, want AMS fonts.
\makeatletter					% Make '@' accessible.
\pagestyle{myheadings}				% We do our own page headers.
\def\@oddhead{\bf Arithmetic Circuits \hfill \myname} % Here they are.
\skip\footins=4ex				% Space above first footnote.
\hbadness=10000					% No "underfull hbox" messages.
\makeatother					% Make '@' special again.
\input{../p.tex}

%==============================================================================
% Title.
%==============================================================================

\begin{document}
\centerline{\LARGE\thishw}

% SECTION: Algebraic Computation Models
\section{Algebraic Computation Models}

Three things to discuss:
\begin{itemize}
\item Algebraic Circuits
\item Computation Trees
\item Blum-Shub-Smale Machines
\end{itemize}

\thm{Definition}{1}{A {\it Field} is:

\begin{itemize}
\item A set $F$
\item The two operators, $+$ and $\times$
\item All elements in the set have an additive and multiplicative inverse, so implicitly $-$ and $\div$ are included as well.
\end{itemize}}

Examples: $\mathbb{R}, \mathbb{C}, \mathbb{Q}$.


\subsection{Algebraic Circuits}

Recall: boolean circuits have $\wedge, \vee, \neg$ gates.

Instead, we now have $+, \times$ gates. Sometimes also allowed the constants $1$ and $-1$, and $\div$.

A circuit is an algebraic circuit if it has one output.

Define polynomials in many variables $f(x_1, \ldots, x_n)$. If $\div$ is allowed, get rational functions $\frac{f}{g}$.


\midline

\elem{Example:} the determinant of a matrix:

$X \in M_n(\mathbb{F})$, $n \times n$ matrix s.t. the elements of $M$ are in the field $\mathbb{F}$.

\begin{equation}
det(x) = \sum_{\sigma \in S_n} (-1)^{sgn(\sigma)} \prod_{i=1}^n x_{i\sigma(i)}
\end{equation}

In general, determinant defined by a polynomial of $n!$ length.

Can compute the determinant via an algebraic circuit of size $\mathcal{O}(n^3)$. Also, there is an $NC^2$ algorithm for computing determinant, gives algorithm form of size $2^{\mathcal{O}(\log^2 n)}$


\elem{Example:} the permanent of a matrix:

Given $X \in M_n(\mathbb{F})$:
\begin{equation}
perm(x) = \sum_{\sigma \in S_n} \prod_{i=1}^n x_{i\sigma(i)}
\end{equation}

There is {\it no known poly size circuit to compute permanent}. Note: it's \#$P$-complete\footnote{Similar to $NP$-complete, but you also count the number of accepting paths in the non-deterministic computation tree}.

\midline

% SECTION: Complexity
\section{Complexity}

Some definitions:

\thm{Definition}{3}{If $\{f_i\}$ is a set of polynomials in $n$ variables of a field $\mathbb{F}$, then this set has poly-bounded degree if there is a $c \in \mathbb{N}$ s.t. each $f_i$ has degree at most $\mathcal{O}(n^c)$.}

\thm{Definition}{4}{The class $AlgP$ is a set of polynomials of poly degree that are computable by algebraic formulas of polynomial size}


\thm{Definition}{5}{$AlgNP$ is the the class of polynomials:
\begin{equation}
f(x_1,x_2,\ldots,x_n) = \sum_{e \in \{0,1\}^{m-n}} g(x_1,\ldots,x_n,e_{n+1}, \ldots, e_m)
\end{equation}
Where $g_n \in AlgP$}

\thm{Definition}{6}{If we have two functions $f(x_1,\ldots,x_n)$ and $g(y_1,\ldots,y_m)$, we say that $f$ is {\it projection reducible} to $g$ if there is a relabeling, $\sigma: \{y_1, \ldots, y_m\} \mapsto \{0,1,x_1,\ldots,x_n\}$ such that: $f(x_1,\ldots,x_n) = g(\sigma(y_1),\ldots,\sigma(y_n))$}

\midline

\thm{Theorem}{1 (Valiant)}{
\begin{enumerate}
\item Every polynomial in $n$ variables computable by circuit of size $u$ is projection reducible to the determinant function on $u+2$ variables.
\item Every function in $AlgNP$ is projection reducible to the permanent function.
\end{enumerate}}

Neat: since we're defined on fields, which are possibly infinite, there is not necessarily a way to create a boolean circuit for each algebraic circuit.

\midline

% SECTION: Blum-Shub-Smale Machines
\section{Blum-Shub-Smale (BSS) Model}

\thm{Definition}{7}{Say we have a field $\mathbb{F}$. A {\it BSS} machine is a Turing Machine in which cells store elements from the field $\mathbb{F}$. Also:
\begin{itemize}
\item Shift state: move left or right
\item Branch state: current cell has value $q_1$, go to cell $q_1$, otherwise go to cell $q_2$.
\item Computation state: replace contents of cell $a$ with $f(a)$, where $f$ is a hardwired rational function , $f = \frac{p}{q}$ for $p,q$ polynomials over $\mathbb{F}$.
\item Register containing an element in the field.
\end{itemize}}


Add other abilities:
\begin{itemize}
\item If we add the ability to compare $a \in \mathbb{F}$, whether $a > 0$, then we can compute anything $P_{poly}$ in polynomial time (and recall that $P_{poly}$ contains undecidable problems).\footnote{Also requires an ordered field, since we're using $>$}
\item If we add $\floor{x}$ then can do integer factorization in poly time. (Shamir)
\end{itemize}

Consider $\mathbb{F} = \mathbb{C}$.

\thm{Definition}{8}{
\begin{enumerate}
\item Then $P_{\mathbb{C}}$ is the set of languages that can be decided by a Turing Machine over $\mathbb{C}$ in polynomial time.
\item $NP_\mathbb{C}$ is the set of languages $L$ s.t. $\exists L_0 \in P_\mathbb{C}$, s.t. $x \in L \equiv \exists(y_1,\ldots, y_{p(n)} \in \mathbb{C}^{p(n)})$ s.t. $(x,y) \in L^0$
\end{enumerate}}


Can also consider 0-1-$NP_{\mathbb{C}} = \{ L \cap \{0,1\}^* \mid L \in NP_\mathbb{C}\}$. And 0-1-$NP_\mathbb{C} \subseteq PSPACE$

\midline


Just like $3SAT$ is the canonical $NP$-complete problem, we have $HN_\mathbb{C}$:

\thm{Definition}{9}{The decision problem $HN_\mathbb{C}$ is: given $p_i$ polynomials in $x_1, \ldots, x_n$, do these polynomials have a common root?}

Note: can convert this into $3SAT$ via: $x \vee y \vee z \leftrightarrow (1-x)(1-y)(1-z) = 0$.

\thm{Theorem}{2}{0-1-$HN_\mathbb{C}$ is complete for 0-1-$NP_\mathbb{C}$.}

\midline

% Subsection: Undecidability
\subsection{Undecidability}

\thm{Definition}{10}{The {\it Mandelbrot Set} is:
\begin{align*}
a \in \mathbb{C}, P_a(z) &= z^2 + a \\
\mathcal{M} &= \{a : P_a(0), P_a(P_a(0)), \ldots, \} \text{ is bounded }
\end{align*}
}

If we have comparison operations, can recognize the complement of $\mathcal{M}$: $a \in \overline{\mathcal{M}} \equiv \exists_jk : |P^k_a(0)| > 2$.

\thm{Theorem}{3}{$\mathcal{M}$ is undecidable}















\end{document}