%==============================================================================
% Notes for CSCI 2951-F (Fall 2015)
%==============================================================================
\newcommand{\myname}{David Abel}

%==============================================================================
% Mini Preamble.
%==============================================================================

\documentclass[11pt]{article} % 10pt article, want AMS fonts.
\makeatletter					% Make '@' accessible.
\pagestyle{myheadings}				% We do our own page headers.
\def\@oddhead{\bf BP \& Circuit Lower Bounds \hfill \myname} % Here they are.
\skip\footins=4ex				% Space above first footnote.
\hbadness=10000					% No "underfull hbox" messages.
\makeatother					% Make '@' special again.
\input{../p.tex}

%==============================================================================
% Title.
%==============================================================================

\begin{document}
\centerline{\LARGE{\bf Complexity Reading Group 10/22/2015}}
\vspace{2mm}
\centerline{\Large {\bf Brown University}}

% --- SECTION: Boolean Circuits  ---
\section{Boolean Circuits}

\thm{Definition}{1}{A {\it circuit} is a directed acyclic graph with $n$ source nodes and $1$ sink node. Each of the remaining nodes are logical gates. The size of the circuit is the number of total nodes}

\thm{Definition}{2}{$P_{poly}$ is the class of languages $L$, decidable by polynomial size circuit families}

Alternatively, can think of $P_{poly}$ as the class of Turing Machines with advice. Suppose:
\begin{equation}
L \in P_{poly} \text{ if } \exists \{\alpha_n\} \text{ with } \alpha_n \in \{0,1\}^{p(n)}, \text{ TM } M, \text{ s.t. } x \in L \iff M(x, \alpha_n) = 1
\end{equation}

% THEOREM
\thm{Theorem}{6.19 (Karp Lipton)}{If $NP \subseteq P_{poly}$, then $PH = \Sigma_2^P$. I.e.:
\begin{equation}
PH = \cup_i \Sigma_i^P = \cup_i \Pi_i^P
\end{equation}}

% PROOF
\elem{Proof:} Note that it is sufficient to show that $\Pi_2^P = \Sigma_2^P$, which we can do by showing $\Pi_2\textsc{SAT} \in \Sigma_2^P$

Recall:
\thm{Definition}{3}{Let $\Pi_2\textsc{SAT}$ be:
\begin{equation}
\phi  \in \Pi_2 SAT \iff \forall_u \exists_v : \phi(u,v) = 1
\end{equation}}

$(\phi, u) \in \Pi_2 SAT \iff \exists v : \phi(u,v) = 1$

Suppose $NP \subseteq P_{poly}$.

Then $\exists$ polynomial size $\{c_n\}$ to solve $\Pi_2 SAT$.

So we can generate a witness using this circuit family in linear time. I.e. $\exists \{c_{n'}\}$ which outputs the witness $c_{n'}(\phi,u) = v$, (where $v$ is the witness).

Can rewrite this statement as:
\begin{equation}
\exists_w : w \in \{0,1\}^{q(n)^2} \forall_u : u \in \{0,1\}^n \phi(u, c_{n'}(\phi, u)) = 1
\label{eq:conc}
\end{equation}

Where $c_{n'}(\phi,u)$ generates the witness for $\phi$.

So what we want to show is that Equation~\ref{eq:conc} is true iff $\phi \in \Pi_2 SAT$

Now, note that Equation~\ref{eq:conc} can be verified in $\Sigma_2^P$. (just look at the quantifiers) \qed

Note: If we show $NP \not \subseteq P_{poly} \rightarrow P \neq NP$

\midline

Q: What are the limitations on this model of computation?

% THEOREM
\thm{Theorem}{6.21}{\begin{equation}
\forall_n : \exists_f : f \{0,1\}^n \rightarrow \{0,1\} : \text{ can't be computed by a circuit $C$ s.t. } |C| = \frac{2^n}{10n}
\end{equation}}

% Proof
\elem{Proof:} Given a circuit of size $\leq S$, it can be represented by $\leq c* S^2$ bits.

If this is our restriction on circuit size, then we can have $s^{3s}$ possible circuits (number of possible DAGs with in-degree 2).

Suppose $S = \frac{2^n}{10n}$, then clearly far fewer circuits than $2^{2^n}$. \qed

I.e. you cannot compute a huge space of functions with polynomial circuits.

\midline

\thm{Theorem}{6.23 (Non-Uniform Hierarchy Theorem)}{

\begin{equation}
\forall_T, T' : \mathbb{N} \rightarrow \mathbb{N} \text{ with } \frac{2^n}{n} > T'(n) > 10T(n) > n
\end{equation}

I.e. $\textsc{Size}(T(n)) \subsetneq \textsc{Size}(T'(n))$

Intuition: With a larger circuit, you can compute strictly more functions.}

\midline

% Subsection: Gate Elimination Method
\subsection{Gate Elimination Method}


Suppose we have a function $f$ that we're trying to compute with a circuit.:
\begin{enumerate}[(1)]
\item Assigning variables in $f$ to maintain properties of $f$.
\item Eliminate gates in our circuit by storing variables, preserve same properties of $f$ via the new circuit.
\end{enumerate}

\thm{Definition}{5}{Let $Q_{2,3}^{(n)}$ be the class of functions where:
\begin{multline}
f : \{0,1\}^n \rightarrow \{0,1\} \in Q_{2,3}^{(n)} \iff \\
\forall_{(x_i, x_j)} \text{ $f$ has at least 3 distinct sub functions as $(x_i,x_j)$ range}
\end{multline}

Also: $\forall x_i : \exists c_i \text{ s.t. } f_{x_i = c_i} \in Q_{2,3}^{(n-1)}$}

Recursion bottoms out at $Q_{2,3}^{(3)}$, since $n = 3$ is the smallest input space for which there can be $3$ distinct sub functions.

\elem{Example:} $f_c^{(n)} (x_1, \ldots, x_n) = \left( \left( \sum_i x_i \right) \text{ mod } 3\right) \text { mod } 2$, for $c \in \{0,1,2\}$.

Now we're going to prove a bound for this class of functions:

% THEOREM
\thm{Theorem}{9.3.2 (from John Hughes' book)}{If $f \in Q_{2,3}^{(n)}, C(f) \geq 2n - 3$, where $C()$ is a circuit.}

\elem{Proof:} 

\underline{Part One}: $f$ depends on each var $x_i$. \qed

\underline{Part Two}: Some input vertex $x_i$ has fan-out $\geq 2$. (suppose we're dealing with a more general class of circuits now). 

Consider a gate $g$ that has the maximum possible length to the output. Since we're in a DAG, $g$ has to be directly receiving input nodes. Suppose $x_i$ and $x_j$ are the variables that feed in to $g$., and they both don't feed anywhere else.

If we fix $x_i$, then there are only two possible sub functions left, which is a contradiction. Therefore, there must be at least one input variable that feeds in to more than one input node (i.e. fan-out $\geq 2$).

By induction: $C(f_{n-1}) \geq 2(n-1) -3$, $C(f_n) \geq 2(n-1) - 3 + 2 \geq 2n - 3$


% --- SECTION: NC and AC ---
\section{NC and AC classes}

\thm{Definition}{6}{$L \in NC^{d}$ if:
\begin{itemize}
\item $L$ can be decided by poly size $\{C_n\}$
\item Depth $\mathcal{O}log^d(n)$
\item Each gate having bounded (2) fan in
\end{itemize}}

\thm{Definition}{7}{$L \in AC^{d}$ if:
\begin{itemize}
\item $L$ can be decided by poly size $\{C_n\}$
\item Depth $\mathcal{O}log^d(n)$
\item Each gate having unbounded fan in
\end{itemize}}

Note: $NC^0 \subsetneq AC^0 \subsetneq NC^1$

And more generally: $NC^i \subseteq AC^i \subseteq NC^{i+1}$

\thm{Theorem}{1 (Ajtai '83)}{ $\textsc{Parity} \not \in AC^0$}



\end{document}